---
title: "allen-electro-morpho"
output: html_document
--- 

```{r}
rm(list = ls())
library(bnlearn)
library(ggplot2)
library(tidyverse)
devtools::load_all()
insertSource("allen-electro-bnlearn-layout.R", 'bnlearn', 
             functions = c('graphviz.backend', 'graphviz.compare.backend', 'graphviz.compare'))
load('../data/allen-electro-morpho.rdata')
```  

# Electro

## Learn bn mouse
```{r} 
db <- ep %>% filter(species == 'Mus musculus') %>% select(-species, -reconstruction_type, -structure_layer_name, -specimen_id)
mouse <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30)
pcorm <- apply(arcs(mouse), 1, function(x) get_pcor(x[1], x[2], db)) %>% round(digits = 2) 
acor <- get_arcs_cor(mouse, db)
acor    %>% arrange(`acor.indi.`)
```    
  

# Learn bn 2/3 human
```{r} 
db <- ep %>% filter(species == 'Homo Sapiens') %>% select(-species, -reconstruction_type, -structure_layer_name, -specimen_id) 
human <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30)  
pcorh <- apply(arcs(human), 1, function(x) get_pcor(x[1], x[2], db)) %>% round(digits = 2)
acor <- get_arcs_cor(human, db)  
acor    %>% arrange(`acor.indi.`)
```

# Compare plot   
```{r}  
pdf(file = 'electro-test.pdf', width = 14, height = 7) 
a <- bnlearn::graphviz.compare(mouse, human, layout = 'fdp', diff = 'none', 
             edgeAttrs=get_edge_attrs(mouse, pcorm))         
mgraph <- format_edges(a[[1]], mouse, morphodb, pcorm)
a <- bnlearn::graphviz.compare(mouse, human, layout = 'fdp', diff = 'none', 
             edgeAttrs=get_edge_attrs(human, pcorh))
             # edgeAttrs=get_edge_attrs(mouse, pcorm))
hgraph <- format_edges(a[[2]], human, morphodb, pcorh) 
dev.off()  
pdf(file = 'electro-mouse.pdf', width = 14, height = 7)  
graph::nodeRenderInfo(mgraph)[["fontsize"]] = 60 
Rgraphviz::renderGraph(mgraph) 
dev.off() 
pdf(file = 'electro-human.pdf', width = 14, height = 7)   
graph::nodeRenderInfo(hgraph)[["fontsize"]] = 60 
Rgraphviz::renderGraph(hgraph) 
dev.off() 
```       

<!-- # Markov blankets Species -->
<!-- - How many features do not have species in their Markov blanket? -->
<!-- - Only "input_resistance" does not have species in MB -->
<!-- - This is symmetric to the other one? -->
<!-- ```{r} -->
<!-- # mbs <- lapply(nodes(orig), function(x) mb(orig, x)) -->
<!-- # spec <- sapply(mbs, function(x) 'species' %in% x) -->
<!-- # nodes(orig)[!spec] -->
<!-- ``` -->

## Human bn
```{r}
marindep <- sapply(nodes(human), function(x) dsep(human, 'rel depth', x))
nodes(human)[!marindep]
mb(human, 'rel depth')
mbsh <- lapply(nodes(human), function(x) mb(human, x))
mbsh <- setNames(mbsh, nodes(human))
sapply(mbsh, length)
```

## Mouse bn
```{r} 
mb(mouse, 'rel depth') 
marindep <- sapply(nodes(mouse), function(x) dsep(mouse, 'rel depth', x))
nodes(human)[!marindep]
mbsm <- lapply(nodes(mouse), function(x) mb(mouse, x))
mbsm <- setNames(mbsm, nodes(mouse)) 
sapply(mbsh, length)
```
  
## Markov blankets across species 
- The Markov blankets differed substantially between mouse and human
- The most similar mb was that of 
```{r}  
mapply(function(a, b) length(intersect(a, b)), mbsh, mbsm)
mapply(function(a, b) length(intersect(a, b)) / length(a), mbsh, mbsm)
mapply(function(a, b) length(intersect(a, b)) / length(b), mbsh, mbsm)
nodes(mouse)[6]
```  
## Compare graphs
```{r}
bnlearn::compare(human, mouse, arcs = TRUE)
```


## Do a test comparison t-test         
- Some of these are actually not significant
```{r}
tvars <- ep %>% select(-species, -reconstruction_type, -structure_layer_name, -specimen_id,
                       -`rel depth`)  %>% mutate(species = ep$species)   
table <- get_ttest_pvals(tvars)  
table[, -1] <- round(table[, -1], 3)
indi <- grep('time', table$variable)
table[-indi, -1] <- round(table[-indi, -1], 2) 
table$Humanm <- table$`mean_Homo Sapiens`
table$Humanpm <- '$\\pm$'
table$Humansd <- table$`sd_Homo Sapiens`
table$Mousem <- table$`mean_Mus musculus`
table$Mousepm <- '$\\pm$'
table$Mousesd <- table$`sd_Mus musculus`
table <- table %>% select(-contains("Musculus"), -contains("Sapiens"))
table <- table[, c('variable', 'Humanm', 'Humanpm', 'Humansd', 'Mousem', 'Mousepm','Mousesd','pvalue')]  
table <- left_join(table, unite) 
table$unit <- paste0('(', table$unit, ')')
table$unit[grep( 'up down', table$variable)] <- ''  
table$variable <- paste0('\\texttt{', table$variable, '} ', table$unit)
table  <- table %>% select(-unit)  
table  <- table %>% rename(`p-value` = pvalue)
label <- 'tbl:ttest'
caption <- 'Per-species mean $\\pm$ standard deviation for each electrophysiological variable, along with the p-value of the t-test.' 
source('../../bbp-interneurons-classify/r/paper.R')
table$variable <-  gsub('_', '\\\\_', table$variable)  
xtable_print(table, label = label,  caption  = caption, include.rownames = FALSE, size = 'normalsize', file = 'ttest.tex')

```
Add this manually into ttest.tex:

      \begin{tabular}{lr@{\hskip 1pt}c@{\hskip 1pt}lr@{\hskip 1pt}c@{\hskip 1pt}lr}
      Variable & \multicolumn{3}{c}{Human} & \multicolumn{3}{c}{Mouse} & p-value \\    

# Morphological   

## Ttest
```{r} 
tvars <- morphodb %>% select(-species, -specimen_id, -`rel depth`)  %>% mutate(species = morphodb$species)  
table <- get_ttest_pvals(tvars)  
table[, -1] <- round(table[, -1], 2)
indi <- match(c('tortuosity', 'diameter', 'angle'), table$variable)
table[-indi, -1] <- round(table[-indi, -1], 0)  
table$Humanm <- table$`mean_Homo Sapiens`
table$Humanpm <- '$\\pm$'
table$Humansd <- table$`sd_Homo Sapiens`
table$Mousem <- table$`mean_Mus musculus`
table$Mousepm <- '$\\pm$'
table$Mousesd <- table$`sd_Mus musculus`   
table <- table %>% select(-contains("Musculus"), -contains("Sapiens")) 
table <- table[, c('variable', 'Humanm', 'Humanpm', 'Humansd', 'Mousem', 'Mousepm','Mousesd','pvalue')]  
table <- left_join(table, unitm)
table$unit <- paste0('(', table$unit, ')')
table$unit[grep( 'tortuosity', table$variable)] <- ''
table$unit[grep( 'order', table$variable)] <- ''
table$variable <- paste0('\\texttt{', table$variable, '} ', table$unit)
table  <- table %>% select(-unit)
label <- 'tbl:ttest-morpho'
caption <- 'Per-species mean $\\pm$ standard deviation for each morphological variable, along with the p-value of the t-test.' 
source('../../bbp-interneurons-classify/r/paper.R')
table$variable <-  gsub('_', '\\\\_', table$variable)   
table  <- table %>% rename(`p-value` = pvalue)
xtable_print(table, label = label,  caption  = caption, include.rownames = FALSE, size = 'normalsize', file = 'ttest-morpho.tex')
```  

Add this manually into ttest-morpho.tex:

      \begin{tabular}{lr@{\hskip 1pt}c@{\hskip 1pt}lr@{\hskip 1pt}c@{\hskip 1pt}lr}
      Variable & \multicolumn{3}{c}{Human} & \multicolumn{3}{c}{Mouse} & p-value \\    

## BN of morpho alone    
```{r}
db <- morphodb %>% filter(species == 'Homo Sapiens') %>% select(-specimen_id, -species)
human <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30)   
pcorh <- apply(arcs(human), 1, function(x) get_pcor(x[1], x[2], db)) %>% round(digits = 2)
cor(db$`rel depth`, db$total_length)
acor <- get_arcs_cor(human, db)
acor    %>% arrange(`acor.indi.`)
```

```{r}
db <- morphodb %>% filter(species == 'Mus musculus') %>% select(-specimen_id, -species)
mouse <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30)   
pcorm <- apply(arcs(mouse), 1, function(x) get_pcor(x[1], x[2], db)) %>% round(digits = 2) 
acor <- get_arcs_cor(mouse, db) 
cor(db$`rel depth`, db$total_length)
cor(db$diameter, db$total_length)
acor    %>% arrange(`acor.indi.`)
```

## Save to pdfs

```{r}  
pdf(file = 'electro-test.pdf', width = 14, height = 7) 
a <- bnlearn::graphviz.compare(mouse, human, layout = 'fdp', diff = 'none', 
             edgeAttrs=get_edge_attrs(mouse, pcorm))         
mgraph <- format_edges(a[[1]], mouse, morphodb, pcorm)
a <- bnlearn::graphviz.compare(mouse, human, layout = 'fdp', diff = 'none', 
             edgeAttrs=get_edge_attrs(human, pcorh))
hgraph <- format_edges(a[[2]], human, morphodb, pcorh) 
dev.off()  
pdf(file = 'morpho-mouse.pdf', width = 14, height = 7)  
graph::nodeRenderInfo(mgraph)[["fontsize"]] = 60 
Rgraphviz::renderGraph(mgraph) 
dev.off() 
pdf(file = 'morpho-human.pdf', width = 14, height = 7)   
graph::nodeRenderInfo(hgraph)[["fontsize"]] = 60 
Rgraphviz::renderGraph(hgraph) 
dev.off()  
```   
## Human MB
```{r} 
mb(human, 'rel depth')
marindep <- sapply(nodes(human), function(x) dsep(human, 'rel depth', x))
nodes(human)[!marindep]
nodes(human)[marindep]
mbsh <- lapply(nodes(human), function(x) mb(human, x))
```


## Mouse MB
```{r}   
mbsm <- lapply(nodes(mouse), function(x) mb(mouse, x))
```

## Compare
```{r} 
mapply(function(a, b) length(intersect(a, b)) / length(a), mbsh, mbsm)
mapply(function(a, b) length(intersect(a, b)) / length(b), mbsh, mbsm)
summary(mapply(function(a, b) length(intersect(a, b)), mbsh, mbsm))
summary(sapply(mbsh, length))
summary(sapply(mbsm, length))
bnlearn::compare(mouse, human, arcs = TRUE  )
# bnlearn::graphviz.compare(mouse, human, layout = 'fdp')  
```
 
# EMorpho
```{r}
m <- morphodb %>% select(-species)
stopifnot(length(unique(m$specimen_id)) == nrow(m) ) 
jep <- ep %>% select(-`rel depth`)
m <- left_join(m, jep, by = 'specimen_id')
stopifnot(length(unique(m$specimen_id)) == nrow(m) )
``` 

- I guess the params cost really little in the Gaussian case, and the BIC is not really good here...
- Again, spurious stuff; the increases should be minimal.
- More arcs with black list in human!
```{r}
db <- m %>% filter(species != 'Mus musculus') %>% select(-reconstruction_type, -structure_layer_name, -specimen_id, -species)  
human <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30, blacklist = get_blacklist(morphodb, ep))   
human <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30)   
pcorh <- apply(arcs(human), 1, function(x) get_pcor(x[1], x[2], db)) %>% round(digits = 2) 
cor(db$tortuosity, db$peak)
cor(db$resistance, db$total_length)
cor(db$resistance, db$width)
acor <- get_arcs_cor(human, db)  %>% mutate(pcor = pcorh)
acor    %>% arrange(`acor.indi.`)
acor %>% filter(to %in% "angle" | par %in% "angle")
```  
   
- For the direction; morpho to electo, make it that way? Use causality.
- It's including some very weak correlations. BIC should be OK with samp size... 
```{r}
db <- m %>% filter(species == 'Mus musculus') %>% select(-reconstruction_type, -structure_layer_name, -specimen_id, -species) 
mouse <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30, blacklist = get_blacklist(morphodb, ep))  
mouse <- tabu(db, score = "bic-g", tabu = 30, max.tabu = 30)
pcorm <- apply(arcs(mouse), 1, function(x) get_pcor(x[1], x[2], db)) %>% round(digits = 2) 
cor(db$resistance, db$distance)
cor(db$resistance, db$width)
cor(db$resistance, db$diameter)
cor(db$resistance, db$tortuosity)
acor <- get_arcs_cor(mouse, db)   %>% mutate(pcor = pcorh)
acor    %>% arrange(`acor.indi.`) 
library(gplots)
rho <- cor(db)
heatmap.2(rho, scale = "none", trace = "none", revC = TRUE)
```

# Plot    
```{r}  
uke <- get_edge_attrs(mouse, pcorm)
uke$label <- uke$label[1:3]  
guke <- get_edge_attrs(human, pcorh)
guke$label <- guke$label[1:3]
```


```{r}  
sizef <- 2.5
pdf(file = 'electro-test.pdf', width = 14 * sizef, height = 7 * sizef) 
a <- bnlearn::graphviz.compare(mouse, human, layout = 'fdp', diff = 'none', 
             edgeAttrs=get_edge_attrs(mouse, pcorm))
             # edgeAttrs=NULL)
             # edgeAttrs=list())
mgraph <- format_edges(a[[1]], mouse, morphodb, pcorm)
a <- bnlearn::graphviz.compare(mouse, human, layout = 'fdp', diff = 'none', 
             edgeAttrs=get_edge_attrs(human, pcorh))
             # edgeAttrs=get_edge_attrs(mouse, pcorm))
             # edgeAttrs=NULL)
             # edgeAttrs=list())
hgraph <- format_edges(a[[2]], human, morphodb, pcorh) 
dev.off()  
pdf(file = 'electrom-mouse.pdf', width = 14 * sizef, height = 7 * sizef) 
graph::nodeRenderInfo(mgraph)[["fontsize"]] = 50 
Rgraphviz::renderGraph(mgraph) 
dev.off() 
pdf(file = 'electrom-human.pdf', width = 14 * sizef, height = 7 * sizef) 
graph::nodeRenderInfo(hgraph)[["fontsize"]] = 50 
Rgraphviz::renderGraph(hgraph) 
dev.off()  
```

```{r}  
em_arcs(human)
em_arcs(mouse)
```

## Human mb
```{r} 
mbsh <- lapply(nodes(human), function(x) mb(human, x))
mbsh <- setNames(mbsh, nodes(human))
sapply(mbsh, length)
```


```{r} 
mbe <- mbsh[get_evars(human, morphodb)] 
mbe <- sapply(mbe, function(x) intersect(x, get_mvars(human, morphodb)))
sapply(mbe, length)
setdiff(get_mvars(human, morphodb), unique(unlist(mbe)))  
```

```{r} 
mbe <- mbsh[get_mvars(human, morphodb)] 
mbe <- sapply(mbe, function(x) intersect(x, get_evars(human, morphodb)))
sapply(mbe, length)
setdiff(get_evars(human, morphodb), unique(unlist(mbe)))  
```

## Mouse MB
s/human/mouse/g
```{r} 
mbsh <- lapply(nodes(mouse), function(x) mb(mouse, x))
mbsh <- setNames(mbsh, nodes(mouse))
sapply(mbsh, length)
```
```{r}
mbsh['resistance']
mbsh['rel depth']
```


```{r} 
mbe <- mbsh[get_evars(mouse, morphodb)] 
mbe <- sapply(mbe, function(x) intersect(x, get_mvars(mouse, morphodb)))
sapply(mbe, length)
setdiff(get_mvars(mouse, morphodb), unique(unlist(mbe)))  
```

## Compare
```{r}
bnlearn::compare(human, mouse, arcs = TRUE)
```
```{r} 
mbsh <- lapply(nodes(human), function(x) mb(human, x))
mbsm <- lapply(nodes(mouse), function(x) mb(mouse, x))
mapply(function(a, b) length(intersect(a, b)), mbsh, mbsm)
a <- mapply(function(a, b) length(intersect(a, b)) / length(a), mbsh, mbsm)
b <- mapply(function(a, b) length(intersect(a, b)) / length(b), mbsh, mbsm)
which(a == 1 & b == 1)
```

